High-accuracy neutron diffusion calculations based on integral transport theory

### 2019-12-01 docs/EPJP2019/response.tex

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{graphicx}% Include figure files
%\usepackage{dcolumn}% Align table columns on decimal point
%\usepackage{bm}% bold math
%\usepackage[sectionbib]{natbib}% bold math
\usepackage{letterbib}
\usepackage[sort&compress]{natbib}% bold math
%\bibliographystyle{abbrv}
\bibliographystyle{unsrt}
%\bibliographystyle{sort}
%\usepackage[english,hebrew]{babel}
\usepackage{hyperref}

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font 
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

\usepackage[usenames, dvipsnames]{color}
% for quotes
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}

% Margins
\topmargin=-1.in % Moves the top of the document 1 inch above the default
\textheight=9.in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=0.2in % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6in % Total width of the text, increase this if the left margin was decreased and vice-versa

%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\newcommand{\tsup}[1]{\textsuperscript{#1}}
\newcommand{\tsub}[1]{\textsubscript{#1}}
\newcommand{\revised}[1]{{\color{red}{#1}}}
\newcommand{\revb}[1]{{\color{magenta}{#1}}}

\newcommand{\keff}{\ensuremath{k_{\texttt{\scriptsize{eff}}}}}
\newcommand{\ppf}{\textrm{PPF}}

\newcommand{\eg}{e.g.,}
\newcommand{\ie}{i.e.,}

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{The European Physical Journal Plus \\
Editorial board}

\signature{
	\vspace{-30pt}
	Erez Gilad, PhD \\
	\mbox{Ben-Gurion University of the Negev}\\
	E-mail: gilade@bgu.ac.il
} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------
\opening{Dear Dr. Jozef Ongena, Managing editor\\}

Thank you for the response regarding the manuscript EPJP-D-19-01168 entitled "High-accuracy neutron diffusion calculations based on integral transport theory". We are also grateful for the insightful comments by the reviewer.

The manuscript underwent a revision in light of the reviewer's comments and is hereby resubmitted for publication in \textit{The European Physical Journal Plus}.

In the revised manuscript we address in full the comments by the reviewer, and the revised parts are marked \revised{in red}. Following are the comments by the reviewer in black followed by our responses (\revised{in red}).

We hope that the Editor and the reviewer find the revised manuscript satisfactory and recommend it for publication.

%--------------------------------------------------------------
% LETTERS ENDS HERE ******************************************
\closing{\mbox{Sincerely yours on behalf of all the co-authors,}}

%---------------------------------------------------------------------------------------	-
\newpage

\textbf{Reviewer \#1}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Please re-read the whole paper to check the language and the punctuation. Some sentences are very concise and could be expanded a little for clarity.}
\\ \\
\revised{The manuscript underwent a thorough examination of the language and punctuation, and some sentences were elaborated for clarity. For example, ...}
\\ \\
% --------------------------------------------------------------------------------------
\textit{I have a concern on which the Authors may want to comment. Formula (3) becomes problematic where the gradient of the flux is small, which is usually the zone where diffusion works better.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{I do not think it would be too hard work to report more results. For instance, I think it would be nice to see the effect of the physical size of the domain, to better detect and evidence the effect of the boundary. This would add a lot to the value of the paper.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{I do not think it would be too hard work to report more results. For instance, I think it would be nice to see the effect of the physical size of the domain, to better detect and evidence the effect of the boundary. This would add a lot to the value of the paper.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In the abstract: I suggest to avoid saying that the results match "extremely well" the reference... while this is true for the eigenvalue, it may be questionable for the flux near the boundary, where effects should be more deeply investigated. Maybe "well" is already enough.}
\\ \\
\revised{T}
\\ \\

\newpage

% --------------------------------------------------------------------------------------
\textit{Page 22, At line 40: largest errors appear near the boundary. This is of course related to the failure of diffusion, however there might be also an additional contribution of the type of boundary conditions adopted (see below). Any comments? In any case, this aspect needs to be investigated.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Line 55: for clarity I would change it to "uncollided neutrons originated from an incoming angular flux at the boundary".}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Page 3, In the title of subsection 2.3 change "\&" with "and".}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In formula (8) it should be specified that it is assumed x'<x, otherwise it is not consistent with Eqs. (7).}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{There is no definition of the moments of the cross sections, Eq. (10).}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Page 5, At line 42 I suggest to change "terms to get" with "terms, one gets".}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{At line 58: the fission operator is not given by q, but, rather, it can be obtained by the observation of the form of the q term, Eq. (25). Since it is crucial to see the effect of the diffusion coefficient update, the operators could be reported explicitly.}
\\ \\
\revised{T}
\\ \\

\newpage

% --------------------------------------------------------------------------------------
\textit{Page 6, To me the title of subsection 3.2 is not correct (there is no such a thing as "integral currents". Should it rather be something like "evaluation of the currents by the integral form of the transport equation"? I suggest also to change the caption of Fig. 2 accordingly.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Since the whole section consider the multigroup formulation, at line 6 to be consistent I suggest to change $J(x,E)$ to $J_g(x)$. }
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Page 7, About boundary conditions: how is the extrapolated distance chosen with vacuum? For instance, with a Mark or a Marshak approach or according to other recipes? And, in any case, what is the value assumed for the calculations? Furthermore, the Authors should also comment on the problem of the need to update the extrapolated distance since it is in general related to the diffusion coefficient (e.g. 2D in Marshak BC). In this respect, the meaning of the sentence in line 53 and 54 is not clear to me.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{page 8, The statement about boundary conditions should be clarified at line 40. What does "they are constant for each energy group" mean? If it means that the same extrapolated distance is assumed for both groups, the Authors should consider to make a more consistent assumption of group-dependent extrapolated distances. The effect of the choice of the BC is not insignificant, especially for optically thin systems. It could be interesting to see the effect of the different choices.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In Table 1 units must be indicated.}
\\ \\
\revised{T}
\\ \\

\newpage

% --------------------------------------------------------------------------------------
\textit{At line 50, please change "w/" with "with". At line 55, please change "w/o" with "without". The use of such contractions (such as "\&" above) is not advisable in a scientific paper. I suggest to make these changes throughout the whole paper.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{At line 60, it is advisable to expand on "the corrected diffusion current exhibit(s)non-trivial behaviour near the boundary". Can it be related to BC?}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{General comment on the Figures: it would be nice to explicitly identify the quantities in the graphs rather than letting the reader struggle to interpret them from the legend.}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In Fig. 8: is there any explanation for the sign change in some curves?}
\\ \\
\revised{T}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Suggestion for further work: it should not be too difficult to study a heterogeneous configuration, since the Ronen method naturally leads to a non-homogeneous system, to see how the Ronen method adapts to interfaces, where strong gradients of the flux might appear.}
\\ \\
\revised{T}
\\ \\

%\bibliography{severe_accident}

\end{letter}

\end{document}

### 2019-12-02 docs/EPJP2019/Gross-EPJP-2019.tex
Update ICTT src files and upload the slides shown at the conference
unstage temporary files and create a directory for figures
Updated tex files for EPJP revision. Work in progress.
The updated ms is Gross-EPJP-2019.tex. The response to reviewer and editor is in new directory response/

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal PLUS
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag / Societa` Italiana di Fisica  %%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{filecontents}{leer.eps}
!PS-Adobe-2.0 EPSF-2.0
%%CreationDate: Mon Jul 13 16:51:17 1992
%%DocumentFonts: (atend)
%%Pages: 0 1
%%BoundingBox: 72 31 601 342
%%EndComments

gsave
72 31 moveto
72 342 lineto
601 342 lineto
601 31 lineto
72 31 lineto
showpage
grestore
%%Trailer
%%DocumentFonts: Helvetica
\end{filecontents}
%
\documentclass[epj]{svjour}
% Remove option referee for final version
%
% Remove any % below to load the required packages
%\usepackage{latexsym}
\usepackage{mathtools} % Loads amsmath
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{xfrac}
\usepackage{graphicx}
%\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref} % for hyperrefs
%\usepackage{cleveref}
%\usepackage[authoryear]{natbib}
\usepackage{gensymb}  % provides macro \degree which works in text and math
\usepackage{color}
%\usepackage{enumitem} 
\usepackage{bm}
\usepackage{multirow}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{arrows}

@@

%
% ---------------------------------------------------
\begin{document}
%
\title{High-accuracy neutron diffusion calculations based on integral transport theory}
%\subtitle{}
\author{
	Roy Gross\inst{1} \and 
	Daniele Tomatis\inst{2} \and 
	Erez Gilad\inst{1}\thanks{\email{gilade@bgu.ac.il} (corresponding author)}
}                     % Do not remove
%
%\offprints{}          % Insert a name or remove this line
%
\institute{
	The Unit of Nuclear Engineering, Ben-Gurion University of the Negev, 8410501 Beer-Sheva, Israel 
	\and
	CEA, DEN, Service d'{\'e}tudes des r\'eacteurs et de math\'ematiques appliqu\'ees (SERMA),
	Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France}
%
\authorrunning{Gross, Tomatis \& Gilad} 
\titlerunning{Integral transport correction to diffusion calculations}
\date{Received: date / Revised version: date}
% The correct dates will be entered by Springer
%
\abstract{
In this paper, the Ronen method is developed, implemented and applied to resolve the neutron flux and the criticality eigenvalue in a simple homogeneous problem. The Ronen method is based on iterative calculations of the multigroup diffusion coefficients using a multigroup diffusion model and driven by the solutions of the integral transport equation. The local diffusion constants are modified in order to reproduce new estimates of the surface currents by a transport operator. The diffusion solver employed in this study uses finite differences and the transport-corrected currents are introduced into the numerical scheme by means of drift terms. The corrected solutions are compared against reference results obtained  by a discrete ordinate code. Boundary conditions are discussed and proper approximations are introduced in order to conserve the particle balance. The results match extremely well the reference solutions, especially in the limit of fine meshes, but slow convergence of the scalar flux is reported.
%
\PACS{
	{28.00.00}{Nuclear engineering and nuclear power studies} \and
	{28.20.Gd}{Neutron transport: diffusion and moderation} \and
	{05.60.-k}{Transport processes} \and
	{28.41.-i}{Fission reactors}
     } % end of PACS codes
  \keywords{
  	neutron diffusion -- integral transport -- non-linear transport correction
  }
} %end of abstract
%
\maketitle
%
% ---------------------------------------------------
\section{Introduction}
\label{sec:intro}

Neutron transport calculations on a full core scale can be a highly intensive computational task. High-fidelity core design optimization or transient analyses can quickly become computationally impractical when using transport methods~\cite{Kim-2019}. For example, in order to achieve a 1\% accuracy in the local flux/power estimation, an order of 10\tsup{11} histories is needed in a full-core Monte Carlo calculation~\cite{Martin-2012}. Another problem is the huge number of variables and parameters to be stored and used during computation, e.g., tallies, geometry, cross sections, and depletion data. A conservative estimation of the memory needed for reasonably accurate full-core Monte Carlo neutronic calculation is in the range of terabytes (TBs)~\cite{Martin-2012}.

%The growing demand for high-accuracy and high-precision full-core computations challenges not only today’s high-end computing systems but will also challenge the near future (e.g., Exascale) computers~\cite{Martin-2012,Smith-2011,Kim-2019}.

@@

In this paper, the development, implementation, and qualification of the Ronen Method~\cite{Ronen-2004} in one-dimensional homogeneous plane geometry is reported. This method is implemented as a highly accurate multigroup neutron diffusion solver based on novel transport corrections. The main hypothesis underlying the Ronen method is based on iterative calculations of the multigroup diffusion coefficients driven by the solutions of the integral transport equation~\cite{Ronen-2004,Tomatis-2011}.

@@

The theoretical background is detailed in~\sxn{sec:theory}, the numerical implementation is described in~\sxn{sec:imp}, the results are presented in~\sxn{sec:res} and conclusions are brought in~\sxn{sec:conc}.

@@

Since these accurate estimates of the currents are based on a known flux distribution, it was also suggested to execute new diffusion calculations, thus updating iteratively the diffusion coefficients in the global calculation. 
\begin{equation}\label{eq:RM-it}
D^{(k+1)}\rE = - \jtr^{(k)}\rE \left[\bnabla\phi^{(k)}\rE\right]^{-1} \,\, ,
\end{equation}
where $k$ is the iteration index. The use of a tensor notation is needed for the diffusion coefficient in general multidimensional problems.

The motivation for this method was to overcome the inherent limitation of Fick's law requiring smooth flux gradients and thus small neutron absorption rate with respect to scattering in general. Nevertheless, isotropic scattering remained as a basic postulate.
For example, in a one-dimensional homogeneous slab with void boundary conditions, Eq.~\eqref{eq:RM-it} takes the following form~\cite{Ronen-2004} 
\begin{equation}\label{eq:RM-it-1D-slab}
D^{(k+1)}(x,E) = -\frac{\frac{1}{2}\int_0^a dx' E_2[\sigma(E)\lvert
	x-x'\rvert]\text{sign}(x-x')q^{(k)}(x',E)}
{\partial \phi^{(k)}(x,E)/\partial x} \,\, .
\end{equation}

@@

There are several ways to drive the integral expression for the flux in slab geometry. Most of them start with assuming homogeneous and isotropic scattering and sources, yielding the \emph{Peierls equation} \cite{Davison-1957,Case-1967,Bell-1970,Pomraning-1973,Duderstadt-1979,Lewis-1984}. The derivation of expression for slab geometry (of thickness $a$) is straightforward, yielding
\begin{eqnarray}\label{eq:flux-slab-l-0-no-bndry-terms}
\phi(x,E)&=&\frac{1}{2}\int_0^a { dx'
	E_1\left[\sigma(E)\lvert x-x'\rvert\right] q(x',E) 
}\nonumber \\
&=& \frac{1}{2}\int_0^a { \!\!\! dx'
	E_1\left[\sigma(E)\lvert x-x'\rvert\right]
	\left[
	\int_0^\infty { \!\!\! dE'
		\sigma_s(E\leftarrow E')\phi(x',E') 
	} + S(x',E)
	\right] 
} \,\, ,
\end{eqnarray}
where $E_1(x)$ is a first order exponential integral~\cite{Abramowitz-1964}. Note that this expression does not include the contribution of uncollided neutrons originate from the flux at the boundaries.

@@

In what follows, a generalization of~\eq{eq:flux-slab-l-0-no-bndry-terms} is derived for heterogeneous medium with anisotropic scattering.  
%
% ---------------------------------------------------
\subsection{Integral expressions for the neutron flux \& current}
\label{sec:int-flux-J}

Consider an infinite one-dimensional slab of width $a$. The angular flux in the slab is given by~\cite{Lewis-1984}
\begin{subequations}\label{eq:int-ang-flx}
	\begin{eqnarray}
	\psi(x,E,\mu) &=& \psi(0,E,\mu) e^{-\mcL(0,x,E)/\mu}
	+ \int_0^x { dx' \,\,
		\frac{Q(x',E,\mu)}{\mu} 
		e^{-\mcL(x',x,E)/\mu} 
	} \,\, , \qquad \mu>0 \,\, ,   \label{eq:int-ang-flx-a}\\
	\psi(x,E,\mu) &=& \psi(a,E,\mu)e^{\mcL(x,a,E)/\mu} 
	-\int_x^a { dx' \,\,
		\frac{Q(x',E,\mu)}{\mu} 
		e^{\mcL(x,x',E)/\mu} 
	} \,\, , \qquad \mu<0 \,\, , \label{eq:int-ang-flx-b}
	\end{eqnarray}
\end{subequations}
where
\begin{equation}\label{eq:opl2}
\mcL(x',x,E) \equiv 
\int_{x'}^x {\sigma(x'',E) dx''} \,\, 
\end{equation}
and $Q(x,E,\mu)$ denotes all sources.

The angular dependence of the source and the uncollided boundary flux terms is expanded in a series of orthogonal Legendre polynomials
\begin{eqnarray}\label{eq:Q-series}
\psi(0,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(0,E) \pl{\mu} 	
} \nonumber \\
\psi(a,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(a,E) \pl{\mu} 	
} \,\, .
\end{eqnarray}
The source cross sections are written explicitly as~\cite{Tomatis-2011}
\begin{equation}\label{eq:src-mmnts}
\sigma_l(x,E\leftarrow E')dE' = 
\sigma_{s,l}(x,E\leftarrow E')dE'
+ \delta_{l0}\frac{\chi(E)}{\keff}\nu\sigma_f(x,E')dE' \,\, 
\end{equation}
and the source term is expanded 
\begin{eqnarray}\label{eq:src}
Q(x,E,\mu)&=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_{s,l}(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} 
+ \frac{\chi(E)}{2\keff} \int_0^\infty { dE'
	\nu\sigma_f(x,E') \psi_0(x,E')
}  \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_l(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
\frac{2l+1}{2} \int_0^\infty { dE'
	q_l(x,E\leftarrow E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} q_l(x,E)\pl{\mu} 			
} 
\,\, ,
\end{eqnarray}
where the source angular moments are defined as
\begin{equation}\label{eq:q_l}
q_l(x,E) \equiv \int_0^\infty dE' q_l(x,E\leftarrow E') \equiv 
\int_0^\infty dE' \sigma_l(x,E\leftarrow E')\psi_l(x,E') dE' \,\, .
\end{equation}

@@

	\draw[>=triangle 45, <->] (7cm,-2.1cm) -- (13cm,-2.1cm);
	\node[anchor=south] at (10cm,-2.1cm) {$\Xi_{i,j}^g + \sigma_{g,j}\left(x'-x_{j-1/2}\right)$};
	\end{tikzpicture}
	\bigbreak
	\caption{Notation for numerical evaluation of integral interface currents.}
	\label{fig:Xi}
\end{figure}

@@

A comparison of the fast and thermal fluxes, as calculated by the reference $S_n$ code, the RM code, and a standard multigroup diffusion w/o RM correction ($D_0$), is shown in Fig.~\ref{fig:fluxes} for half-slab. The deviations (in [\%]) of the RM-corrected flux and the standard diffusion ($D_0$) flux from the reference $S_n$ flux is also shown in Fig.~\ref{fig:fluxes}. The deviation from the reference solution is decreased by the Ronen iterations from approximately 20\% near the boundary and 1\% at the slab center to 2\% and 0\%, respectively.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal PLUS
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag / Societa` Italiana di Fisica  %%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{filecontents}{leer.eps}
%!PS-Adobe-2.0 EPSF-2.0
%%%CreationDate: Mon Jul 13 16:51:17 1992
%%%DocumentFonts: (atend)
%%%Pages: 0 1
%%%BoundingBox: 72 31 601 342
%%%EndComments
%
%gsave
%72 31 moveto
%72 342 lineto
%601 342 lineto
%601 31 lineto
%72 31 lineto
%showpage
%grestore
%%%Trailer
%%%DocumentFonts: Helvetica
%\end{filecontents}
%
\documentclass[epj]{svjour}
% Remove option referee for final version
%
% Remove any % below to load the required packages
%\usepackage{latexsym}
\usepackage{mathtools} % Loads amsmath
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{xfrac}
\usepackage{graphicx}
%\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref} % for hyperrefs
%\usepackage{cleveref}
%\usepackage[authoryear]{natbib}
\usepackage{gensymb}  % provides macro \degree which works in text and math
\usepackage{color}
%\usepackage{enumitem} 
\usepackage{bm}
\usepackage{multirow}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{arrows}

@@

%
\graphicspath{ {./figures/} }
% ---------------------------------------------------
\begin{document}
%
\title{High-accuracy neutron diffusion calculations based on integral transport theory}
%\subtitle{}
\author{
	Roy Gross\inst{1} \and 
	Daniele Tomatis\inst{2} \and 
	Erez Gilad\inst{1}\thanks{\email{gilade@bgu.ac.il} (corresponding author)}
}                     % Do not remove
%
%\offprints{}          % Insert a name or remove this line
%
\institute{
	The Unit of Nuclear Engineering, Ben-Gurion University of the Negev, 8410501 Beer-Sheva, Israel 
	\and
	CEA, DEN, Service d'{\'e}tudes des r\'eacteurs et de math\'ematiques appliqu\'ees (SERMA),
	Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France}
%
\authorrunning{Gross, Tomatis \& Gilad} 
\titlerunning{Integral transport correction to diffusion calculations}
\date{Received: date / Revised version: date}
% The correct dates will be entered by Springer
%
\abstract{
In this paper, the Ronen method is developed, implemented, and applied to resolve the neutron flux and the criticality eigenvalue \revised{in one-dimensional simple homogeneous problem and in more complex heterogeneous benchmarks}. The Ronen method is based on iterative calculations of the multigroup diffusion coefficients using a multigroup diffusion model and \revised{is} driven by the solutions of the integral transport equation. The \revised{spatially-dependent} diffusion constants are modified \revised{locally} in order to produce new estimates of the surface currents \revised{using an integral transport expression}. The diffusion solver employed in this study uses finite differences and the transport-corrected currents are introduced into the numerical scheme \revised{using} drift terms. The corrected solutions are compared against reference results obtained by a discrete ordinate code. Boundary conditions are discussed and proper approximations are introduced to conserve the particle balance. The results match well with the reference solutions, especially in the limit of fine meshes, but slow convergence of the scalar flux is reported.
% 
\PACS{
	{28.00.00}{Nuclear engineering and nuclear power studies} \and
	{28.20.Gd}{Neutron transport: diffusion and moderation} \and
	{05.60.-k}{Transport processes} \and
	{28.41.-i}{Fission reactors}
     } % end of PACS codes
  \keywords{
  	neutron diffusion -- integral transport -- non-linear transport correction
  }
} %end of abstract
%
\maketitle
%
% ---------------------------------------------------
\section{Introduction}
\label{sec:intro}

Neutron transport calculations on a full core scale can be a highly intensive computational task. High-fidelity core design optimization or transient analyses can quickly become computationally impractical when using transport methods~\cite{Kim-2019}. For example, to achieve a 1\% accuracy in the local flux/power estimation, an order of 10\tsup{11} histories is needed in a full-core Monte Carlo calculation~\cite{Martin-2012}. Another problem is the huge number of variables and parameters to be stored and used during computation, e.g., tallies, geometry, cross sections, and depletion data. A conservative estimation of the memory needed for reasonably accurate full-core Monte Carlo neutronic calculation is in the range of terabytes (TBs)~\cite{Martin-2012}. \revised{The growing demand for high-accuracy and high-precision full-core computations challenges not only today’s high-end computing systems but will also challenge the near future (e.g., Exascale) computers~\cite{Martin-2012,Smith-2011,Kim-2019}.}

@@

In this paper, the development, implementation, and qualification of the Ronen Method~\cite{Ronen-2004} in one-dimensional homogeneous \revised{and heterogeneous} plane geometry is reported. This method is implemented as a highly accurate multigroup neutron diffusion solver based on novel transport corrections. The main hypothesis underlying the Ronen method is based on iterative calculations of the multigroup diffusion coefficients driven by the solutions of the integral transport equation~\cite{Ronen-2004,Tomatis-2011}.

@@

The theoretical background is detailed in~\sxn{sec:theory}, the numerical implementation is described in~\sxn{sec:imp}, the results are presented in~\sxn{sec:res}, and conclusions are brought in~\sxn{sec:conc}.

@@

Since these accurate estimates of the currents are based on a known flux distribution, it was also suggested to execute new diffusion calculations, thus updating iteratively the diffusion coefficients in the global calculation. 
\begin{equation}\label{eq:RM-it}
D^{(k+1)}\rE = -\frac{\lvert\jtr^{(k)}\rE\rvert}{\lvert\bnabla\phi^{(k)}\rE\rvert} \,\, ,
\end{equation}
where $k$ is the iteration index. \revised{The use of a tensor notation is needed for the diffusion coefficient in general multidimensional problems.}

The motivation for this method was to overcome the inherent limitation of Fick's law requiring smooth flux gradients and thus small neutron absorption rate with respect to scattering in general. Nevertheless, isotropic scattering remained as a basic postulate.
For example, in a homogeneous slab with void boundary conditions, Eq.~\eqref{eq:RM-it} takes the following form~\cite{Ronen-2004} 
\begin{equation}\label{eq:RM-it-1D-slab}
D^{(k+1)}(x,E) = -\frac{\frac{1}{2}\int_0^a dx' E_2[\sigma(E)\lvert
	x-x'\rvert]\text{sign}(x-x')q^{(k)}(x',E)}
{\partial \phi^{(k)}(x,E)/\partial x} \,\, .
\end{equation}

@@

There are several ways to drive the integral expression for the flux in slab geometry. Most of them start with assuming homogeneous and isotropic scattering and sources, yielding the \emph{Peierls equation} \cite{Davison-1957,Case-1967,Bell-1970,Pomraning-1973,Duderstadt-1979,Lewis-1984}. The derivation of expression for slab geometry (of thickness $a$) is straightforward, yielding
\begin{eqnarray}\label{eq:flux-slab-l-0-no-bndry-terms}
\phi(x,E)&=&\frac{1}{2}\int_0^a { dx'
	E_1\left[\sigma(E)\lvert x-x'\rvert\right] q(x',E) 
}\nonumber \\
&=& \frac{1}{2}\int_0^a { \!\!\! dx'
	E_1\left[\sigma(E)\lvert x-x'\rvert\right]
	\left[
	\int_0^\infty { \!\!\! dE'
		\sigma_s(E\leftarrow E')\phi(x',E') 
	} + S(x',E)
	\right] 
} \,\, ,
\end{eqnarray}

where $E_1(x)$ is a first order exponential integral~\cite{Abramowitz-1964}. Note that this expression does not include the contribution of uncollided neutrons \revised{originated from an incoming angular} flux at the boundaries.

@@

In what follows, a generalization of~\eq{eq:flux-slab-l-0-no-bndry-terms} is derived for heterogeneous medium with anisotropic scattering.  
%
% ---------------------------------------------------
\subsection{Integral expressions for the neutron flux \revised{and} current}
\label{sec:int-flux-J}

Consider an infinite slab of width $a$. The angular flux in the slab is given by~\cite{Lewis-1984}
\begin{subequations}\label{eq:int-ang-flx}
	\begin{eqnarray}
	\psi(x,E,\mu) &=& \psi(0,E,\mu) e^{-\mcL(0,x,E)/\mu}
	+ \int_0^x { dx' \,\,
		\frac{Q(x',E,\mu)}{\mu} 
		e^{-\mcL(x',x,E)/\mu} 
	} \,\, , \qquad \mu>0 \,\, ,   \label{eq:int-ang-flx-a}\\
	\psi(x,E,\mu) &=& \psi(a,E,\mu)e^{\mcL(x,a,E)/\mu} 
	-\int_x^a { dx' \,\,
		\frac{Q(x',E,\mu)}{\mu} 
		e^{\mcL(x,x',E)/\mu} 
	} \,\, , \qquad \mu<0 \,\, , \label{eq:int-ang-flx-b}
	\end{eqnarray}
\end{subequations}
where
\begin{equation}\label{eq:opl2}
\mcL(x',x,E) \equiv 
\int_{x'}^x {\sigma(x'',E) dx''} \,\,, \quad \text{assuming }(x'<x)
\end{equation}
and $Q(x,E,\mu)$ denotes all sources.

The angular dependence of the source and the uncollided boundary flux terms is expanded in a series of orthogonal Legendre polynomials
\begin{eqnarray}\label{eq:Q-series}
\psi(0,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(0,E) \pl{\mu} 	
} \nonumber \\
\psi(a,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(a,E) \pl{\mu} 	
} \,\, .
\end{eqnarray}
The source cross sections are written explicitly as~\cite{Tomatis-2011}
\begin{equation}\label{eq:src-mmnts}
\sigma_l(x,E\leftarrow E')dE' = 
\sigma_{s,l}(x,E\leftarrow E')dE'
+ \delta_{l0}\frac{\chi(E)}{\keff}\nu\sigma_f(x,E')dE' \,\, ,
\end{equation}
\revised{where
\begin{equation}\label{eq:src-mmnts-explicit}
\sigma_{s,l}(x,E\leftarrow E') = \int_{-1}^1 d\mu' P_l(\mu')\sigma_s(x,E\leftarrow E',\mu_0)
\end{equation}
with $\mu_0=\OO'\cdot\OO$} and the source term is expanded 
\begin{eqnarray}\label{eq:src}
Q(x,E,\mu)&=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_{s,l}(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} 
+ \frac{\chi(E)}{2\keff} \int_0^\infty { dE'
	\nu\sigma_f(x,E') \psi_0(x,E')
}  \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_l(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
\frac{2l+1}{2} \int_0^\infty { dE'
	q_l(x,E\leftarrow E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} q_l(x,E)\pl{\mu} 			
} 
\,\, ,
\end{eqnarray}
where the source angular moments are defined as
\begin{equation}\label{eq:q_l}
q_l(x,E) \equiv \int_0^\infty dE' q_l(x,E\leftarrow E') \equiv 
\int_0^\infty dE' \sigma_l(x,E\leftarrow E')\psi_l(x,E') dE' \,\, .
\end{equation}

@@

	\draw[>=triangle 45, <->] (7cm,-2.1cm) -- (13cm,-2.1cm);
	\node[anchor=south] at (10cm,-2.1cm) {$\Xi_{i,j}^g + \sigma_{g,j}\left(x'-x_{j-1/2}\right)$};
	\end{tikzpicture}
	\bigbreak
	\caption{\revised{Notation for numerical evaluation of interface currents by the integral form of the transport equation.}}
	\label{fig:Xi}
\end{figure}

@@

A comparison of the fast and thermal fluxes, as calculated by the reference $S_n$ code, the RM code, and a standard multigroup diffusion without RM correction ($D_0$), is shown in Fig.~\ref{fig:fluxes} for half-slab. The deviations (in [\%]) of the RM-corrected flux and the standard diffusion ($D_0$) flux from the reference $S_n$ flux is also shown in Fig.~\ref{fig:fluxes}. The deviation from the reference solution is decreased by the Ronen iterations from approximately 20\% near the boundary and 1\% at the slab center to 2\% and 0\%, respectively.

@@

%
% ---------------------------------------------------
\subsection{Heterogeneous case}
\label{subsec:heterog}
